LG5 Function Errata Updated

### 2020-11-16 doc/LG5_errata/LG5_errata_updated.Rmd
Proof of error with doy_ip

---
title: "LG5 Function Errata"
author: "Albert Y. Kim"
date: "Last updated on `r Sys.Date()`"
output:
  pdf_document:
    number_sections: yes
  html_document:
    df_print: kable
    highlight: tango
    number_sections: yes
    theme: cosmo
urlcolor: blue
---

@@

# $r$ parameter from LG5 model

I have a suspicion both these interpretations of the $r$ parameter of LG5 model are incorrect:

1. McMahon and Parker 2015 state just above Eq 1 (page 245): "the rate parameter $r$ describes the slope of the curve at the inflection point"
1. D'Orangeville et al. state (page 9): "$r$ represents the maximum growth rate" and "$doy_{ip}$ is the day of year when maximum growth rate occurs"

Rather, I believe that the correct formulation of the slope/the maximum growth rate is instead:

@@

Let's set arbitrarily chosen parameter values then

@@

```{r, echo = TRUE}
L <- 13
K <- 14
doy.ip <- 175
r <- 0.05
theta <- 1
```

We then plot both the diameter curve as well as the growth curve, with `doy.ip` marked with a vertical dotted line. Observe how growth peaks at `doy.ip`.

@@

@@

1. In the left plot how the orange line does not appear to be tangent to the LG5 curve at `doy.ip` and hence is not the slope at that point.
1. In the right plot $r$ = `r r` does not correspond to the growth rate at `doy.ip` which should be the max growth rate. Rather it appears to be around 0.012

## Proposed new $r$

Let's repeat the above with our proposed growth rate $r_{new}$, which is $\frac{d}{dx}LG5(x)$ evaluated at $x$ = `doy.ip`. See this [PDF](https://drive.google.com/file/d/1CMT2o1svTfbS7wQ_uJoWjqHbvDopkCZw/view?usp=sharing) for the analytic derivation of the derivative.

@@

```{r, fig.width=10}
# Comupte new r, then find intercept
r_new <- ((K-L)/(1 + 1/theta)^2) * (r/theta)
intercept_new <- y - r_new*doy.ip

par(mfrow=c(1, 2))
plot(dbh_growth_values$doy, dbh_growth_values$diameter, type = "l", 
     main = "LG5 diameter", xlab = "doy", ylab = "diameter")
abline(v = doy.ip, lty = "dotted")
abline(intercept_new, r_new, col = "orange")
plot(dbh_growth_values$doy, dbh_growth_values$growth, type = "l", 
     main = "Growth curve", xlab = "doy", ylab = "growth")
abline(v = doy.ip, lty = "dotted")
abline(h = r_new, col = "orange")
```

1. In the left plot how the orange line is now tangent to the LG5 curve at `doy.ip` and thus is the correct slope.
1. In the right plot $r_{new}$ = `r r_new` now corresponds to the max growth rate occurring at `doy.ip`.

>>>>

---
title: "LG5 Function Errata Updated"
author: "Albert Y. Kim"
date: "Last updated on `r Sys.Date()`"
output:
  pdf_document:
    number_sections: yes
  html_document:
    df_print: kable
    highlight: tango
    number_sections: yes
    theme: cosmo
urlcolor: blue
---

@@

# $doy_{ip}$ and $r$ parameter from LG5 model

I have a suspicion both these interpretations of the $r$ and $doy_{ip}$ parameters of LG5 model are incorrect:

1. McMahon and Parker 2015 state just above Eq 1 (page 245): "$doy_{ip}$ marks the day of the year when the inflection of the curve is predicted to occur, and the rate parameter $r$ describes the slope of the curve at the inflection point"
1. D'Orangeville et al. state (page 9): "$r$ represents the maximum growth rate" and "$doy_{ip}$ is the day of year when maximum growth rate occurs" and "$doy_{ip}$ is the day of year (doy) when maximum growth rate occurs..."

Rather, I believe that the correct formulations of (1) the slope/the maximum growth rate and (2) inflection points are instead:

$$
\begin{aligned}
doy_{ip}^* &= -\frac{\log(\theta)}{r} + doy_{ip}\\
r^* &= \frac{r(K-L)}{4}
\end{aligned}
$$

@@

Let's set arbitrarily chosen parameter values **but making sure to use a value other than $\theta$ = 1**:

@@

```{r, echo = TRUE}
L <- 13
K <- 14
doy.ip <- 175
r <- 0.05
# Use a value other than theta = 1:
theta <- 1.5
```

## Original interpretations of $doy_{ip}$

We then plot both the diameter curve as well as the growth curve, with `doy.ip` marked with a vertical dotted line.

@@

Observe how `doy.ip` is not the day the peak growth occurs.

@@

1. In the left plot how the orange line does not appear to be tangent to the LG5 curve at `doy.ip` and hence is not the slope at that point.
1. In the right plot $r$ = `r r` does not correspond to the growth rate at the peak of the curve, rather it appears to be around 0.012.

## Proposed $doy_{ip}^*$ and $r^*$

Let's repeat the above with our proposed $doy_{ip}^*$, which is the point $x$ at which $\frac{d^2}{dx^2}LG5(x) = 0$ (per calculus on inflection points), and our proposed new growth rate $r^*$, which is $\frac{d}{dx}LG5(x)$ evaluated at $x$ = $doy_{ip}^*$. See [this](https://drive.google.com/file/d/1CMT2o1svTfbS7wQ_uJoWjqHbvDopkCZw/view?usp=sharing) and [this](https://drive.google.com/file/d/10MjbohLvOvLxkBpNmC_bnJKnLiE0K-pr/view?usp=sharing) PDF for all analytic derivative derivations.

$$
\begin{aligned}
r^* &= \frac{r(K-L)}{4}\\
doy_{ip}^* &= -\frac{\log(\theta)}{r} + doy_{ip}
\end{aligned}
$$

```{r, fig.width=10}
# Comupte new r and doy.ip, then find intercept
r_new <- r*(K-L)/4
doy.ip_new <- -log(theta)/r + doy.ip
y <- lg5.pred(params, doy.ip_new)
intercept_new <- y - r_new*doy.ip_new

par(mfrow=c(1, 2))
plot(dbh_growth_values$doy, dbh_growth_values$diameter, type = "l", 
     main = "LG5 diameter", xlab = "doy", ylab = "diameter")
abline(v = doy.ip_new, lty = "dotted")
abline(intercept_new, r_new, col = "orange")
plot(dbh_growth_values$doy, dbh_growth_values$growth, type = "l", 
     main = "Growth curve", xlab = "doy", ylab = "growth")
abline(v = doy.ip_new, lty = "dotted")
abline(h = r_new, col = "orange")
```

1. In the left plot how the orange line is now tangent to the LG5 curve at `doy.ip` and thus is the correct slope.
1. In the right plot $r^*$ = `r r_new` now corresponds to the max growth rate correctly occurring at `doy.ip`.

# Previous errata

In the previous errata PDF `LG5_errata.pdf` I suggested that the correct max growth rate was

@@

However, this result was based on the incorrect assumption that $doy_{ip}$ was the correct inflection point, which we evaluated $\frac{d}{dx}LG5(x)$ to obtain the max growth rate. Furthermore, the empirical simulations/plots I showed were for $\theta = 1$. If we plug this value into the previous $r_{new}$ we get

$$
r_{new} = \frac{K-L}{\left(1 + \frac{1}{\theta}\right)^2}\frac{r}{\theta} = \frac{K-L}{\left(1 + \frac{1}{1}\right)^2}\frac{r}{1} = \frac{r(K-L)}{4}
$$
which is our updated formulation above.
